---
title: "Statistical Inference Course Project (Part 1)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: paged
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

## Overview

The Central Limit Theorem states that if you have a population with mean $\mu$
and standard deviation $\sigma$ and take sufficiently large random samples from
the population (generally sample sizes greater than 30), then the distribution of the
sample means will be approximately normally distributed about the population
mean $\mu$ - no matter the shape of the population distribution.

This project explores the Central Limit Theorem using the exponential
distribution in R. The theoretical normal distribution will be compared to the
distribution of calculated means of samples from the exponential distribution.

## Simulations

Perform 1000 simulations, each with 40 samples of an exponential distribution.
The 40 samples will be used to calculate the arithmetic mean and variance and
then compared to the theoretical estimates.

To make the data reproducible, a seed will be set. Also, set the control 
parameters $\lambda = 0.2$ (the rate) and $n = 40$ (number of samples). A histogram
will be provided to show the averages of the 40 exponentials over 1000
simulations.

```{r exp-simulation, echo = TRUE}
# load libraries
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}

# set seed for reproducability
set.seed(006766)

# set sampling values:
lambda <- 0.2             # rate parameter
n <- 40                   # number of samples (exponentials) in each simulation
numSimulations <- 1000    # number of simulations

# simulate the population
simMeans <- data.frame(expMean = sapply(1 : numSimulations, function(x) {mean(rexp(n, lambda))}))

# plot the distribution
expSimulationMeansChart <- ggplot(simMeans, aes(x = expMean, y = ..count..)) +
                                  geom_histogram(binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
                                  xlab("Mean") +
                                  ylab("Frequency") +
                                  theme(plot.title = element_text(size = 14, hjust = 0.5)) +
                                  ggtitle("Histogram of Exponential Simulation Means")
print(expSimulationMeansChart)
```

## Sample Mean versus Theoretical Mean

*Pending*

The mean of the exponential distribution is `1/lambda` and the standard
deviation is also `1/lambda`.



Perform 1000 simulations on 40 samples which are exponentially distributed
with `lambda = 0.2`, then take the mean of each simulation.

In this project, investigate the exponential distribution in R and compare it
with the Central Limit Theorem (CLM). The exponential distribution will be
simulated in R with `rexp(n, lambda)` where lambda is the rate parameter.

Lambda will be set to `lambda = 0.2` for all of
the simulations. An investigation will be performed on the distribution via
simulation of averages of 40 exponentials for one thousand simulations.



sampleMean <- mean(simMeans$mValue)
theoMean <- 1/lambda
print(sampleMean)
print(theoMean)

## Sample Variance versus Theoretical Variance



## Distribution
