---
title: "Statistical Inference Course Project (Part 1)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: paged
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

## Overview

The Central Limit Theorem states that if you have a population with mean $\mu$
and standard deviation $\sigma$ and take sufficiently large random samples from
the population (generally sample sizes greater than 30), then the distribution
of the sample means will be approximately normally distributed about the
population mean $\mu$ - no matter the shape of the population distribution.

This project explores the Central Limit Theorem using the exponential
distribution in R. The theoretical normal distribution will be compared to the
distribution of calculated means of samples from the exponential distribution.

## Simulations

Perform 1000 simulations, each with 40 samples of an exponential distribution.
The 40 samples will be used to calculate the arithmetic mean and variance and
then compared to the theoretical estimates.

To make the data reproducible, a seed will be set. Also, set the control 
parameters $\lambda = 0.2$ (the rate) and $n = 40$ (number of samples). A histogram
will be provided to show the averages of the 40 exponentials over 1000
simulations.

```{r exp-simulation, echo = TRUE}
# load libraries
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}

# set seed for reproducability
set.seed(062000)

# set sampling values:
lambda <- 0.2             # rate parameter
n <- 40                   # number of samples (exponentials) in each simulation
numSimulations <- 1000    # number of simulations

# simulate the population
simMeans <- data.frame(expMean = sapply(1 : numSimulations, function(x) {mean(rexp(n, lambda))}))

# plot the distribution
expSimulationMeansChart <- ggplot(simMeans, aes(x = expMean, y = ..count..)) +
    geom_histogram(binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
    xlab("Mean") +
    ylab("Frequency") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Histogram of Exponential Simulation Means")
print(expSimulationMeansChart)
```

## Sample Mean versus Theoretical Mean

According to the Central Limit Theorem, the distribution of the sample means
will be approximately normally distributed with a mean equal to the population
mean $\mu$ of the underlying distribution. Because the underlying distribution
in this simulation is exponential, the theoretical mean of the exponential
distribution will be compared to the corresponding sample mean of the
simulation. For an exponential distribution, the theoretical mean is equal to
$\frac{1}{\lambda}$. 

Calculate the sample mean and theoretical mean across all 1000 simulations of
40 samples from an exponential distribution where $\lambda = 0.2$. 

```{r sample-vs-theoretical-mean, echo = TRUE}
# calculate sample mean and theoretical mean
sampleMean <- mean(simMeans$expMean)
theoMean <- 1/lambda
compMeans <- data.frame(sampleMean, theoMean)
names(compMeans) <- c("Sample Mean", "Theoretical Mean")
print(compMeans)

# plot the distribution (sample mean versus theoretical mean)
expSimulationMeansChart <- ggplot(simMeans, aes(x = expMean, y = ..count..)) +
    geom_histogram(binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
    geom_vline(xintercept = sampleMean, color = "blue", size = 0.25) +
    geom_vline(xintercept = theoMean, color = "red", size = 0.25) +
    xlab("Mean") +
    ylab("Frequency") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Histogram of Exponential Simulation Means")
print(expSimulationMeansChart)
```

The sample mean came out to be `r sampleMean` while the theoretical mean is
`r theoMean`. As shown in the above chart, the mean of the sample means of
exponentials (blue vertical line) is very close to the theoretical mean of an
exponential distribution (red vertical line).

## Sample Variance versus Theoretical Variance

and a standard deviation equal to the population standard deviation of the underlying distribution divided by the square root of the sample size ($\frac{\sigma}{\sqrt{n}}$).

## Distribution
