---
title: "Statistical Inference Course Project (Part 1)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

## Overview

The Central Limit Theorem states that if you have a population with mean $\mu$
and standard deviation $\sigma$ and take sufficiently large random samples from
the population (generally sample sizes over 30), then the distribution of the
sample means will be approximately normally distributed about the population
mean $\mu$ - no matter the shape of the population distribution.

This project explores the Central Limit Theorem using the exponential
distribution in R. The theoretical normal distribution will be compared to the
distribution of calculated means of samples from the exponential distribution.

## Simulations

*Pending*

Perform 1000 simulations on 40 samples which are exponentially distributed
with `lambda = 0.2`, then take the mean of each simulation.

In this project, investigate the exponential distribution in R and compare it
with the Central Limit Theorem (CLM). The exponential distribution will be
simulated in R with `rexp(n, lambda)` where lambda is the rate parameter.

The mean of the exponential distribution is `1/lambda` and the standard
deviation is also `1/lambda`. Lambda will be set to `lambda = 0.2` for all of
the simulations. An investigation will be performed on the distribution via
simulation of averages of 40 exponentials for one thousand simulations.







```{r exp-simulation, echo = TRUE}
# load libraries
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}

# set seed for reproducability
set.seed(006867)

# set sampling values:
lambda <- 0.2             # rate parameter
n <- 40                   # number of samples (exponentials) in each simulation
numSimulations <- 1000    # number of simulations

# simulate the population
simMeans <- data.frame(mValue = sapply(1 : numSimulations, function(x) {mean(rexp(n, lambda))}))

# plot the distribution
expSimulationMeansChart <- ggplot(simMeans, aes(x = mValue)) +
                                  geom_histogram(binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
                                  xlab("Mean") +
                                  ylab("Frequency") +
                                  theme(plot.title = element_text(size = 14, hjust = 0.5)) +
                                  ggtitle("Histogram of Exponential Simulation Means")
print(expSimulationMeansChart)

```

## Sample Mean vs Theoretical Mean

sampleMean <- mean(simMeans$mValue)
theoMean <- 1/lambda
print(sampleMean)
print(theoMean)

## Sample Variance vs Theoretical Variance



## Distribution
