---
title: "Statistical Inference Course Project (Part 1)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

## Overview

The Central Limit Theorem states that if you have a population with mean $\mu$
and standard deviation $\sigma$ and take sufficiently large random samples from
the population (generally sample sizes over 30), then the distribution of the
sample means will be approximately normally distributed about the population
mean $\mu$ - no matter the shape of the population distribution.

In this project, investigate the exponential distribution in R and compare it
with the Central Limit Theorem (CLM). The exponential distribution will be
simulated in R with `rexp(n, lambda)` where lambda is the rate parameter.

The mean of the exponential distribution is `1/lambda` and the standard
deviation is also `1/lambda`. Lambda will be set to `lambda = 0.2` for all of
the simulations. An investigation will be performed on the distribution via
simulation of averages of 40 exponentials for one thousand simulations.

## Simulations

```{r exp-simulation, echo = TRUE}
# load libraries
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}

# set seed for reproducability
set.seed(006867)

# set sampling values:
lambda <- 0.2             # rate parameter
n <- 40                   # number of samples (exponentials) in each simulation
numSimulations <- 1000    # number of simulations

# simulate the population by appending each sample mean to a vector
#simMeans = NULL
#for (i in 1 : numSimulations) {
#    simSample <- rexp(n, lambda)
#    simMeans <- c(simMeans, mean(simSample))
#}
#sampleMean <- mean(simMeans)
#theoMean <- 1/lambda

simMeans <- data.frame(x=sapply(1 : 1000, function(x) {mean(rexp(n, lambda))}))
head(simMeans)

#print(sampleMean)
#print(theoMean)

# plot the means distribution

#hist.pop <- ggplot(simMeans, aes(x = x)) + 
#  geom_histogram(aes(y=..count.., fill=..count..)) +
#  labs(title="Histogram for Averages of 40 Exponentials over 1000 Simulations", y="Frequency", x="Mean")
#hist.pop

expSimulationMeansChart <- ggplot(simMeans, aes(x = x)) +
                                  geom_histogram(binwidth = 0.25, aes(y = ..count.., fill = ..count..)) +
                                  xlab("Mean") +
                                  ylab("Frequency") +
                                  theme(plot.title = element_text(size = 14, hjust = 0.5)) +
                                  ggtitle("Exponential Simulation Means")
print(expSimulationMeansChart)

```

## Sample Mean versus Theoretical Mean



## Sample Variance versus Theoretical Variance



## Distribution
