---
title: "Statistical Inference Course Project (Part 1)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
  word_document:
    toc: yes
    df_print: paged
    keep_md: no
---

## Course Project

**Statistical Inference Course Project**

Peer-graded Assignment

* This course project is available on GitHub

    <a href="https://github.com/oraclejavanet/statistical-inference-course-project-1" target="_top">Statistical Inference Course Project</a>

## Synopsis

The Central Limit Theorem states that if you have a population with mean $\mu$
and standard deviation $\sigma$ and take sufficiently large random samples from
the population (generally sample sizes greater than 30), then the distribution
of the sample means will be approximately normally distributed about the
population mean $\mu$ - no matter the shape of the population distribution.

This project explores the Central Limit Theorem using the exponential
distribution in R. The theoretical normal distribution will be compared to the
distribution of calculated means of samples from the exponential distribution.

## Environment Setup

```{r setup, include = FALSE}
# set knitr options
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')

# clear global workspace (including hidden objects) for this session
rm(list = ls(all.names = TRUE))

# free up memory and display statistics on free memory
gc()

# disable scientific notation for numbers
options(scipen = 1)
```

Load packages used in this analysis.

```{r load-packages, echo = TRUE}
if (!require(ggplot2)) {
    install.packages("ggplot2")
    library(ggplot2)
}
```

Display session information.

```{r display-session-info, echo = TRUE}
sessionInfo()
```

## Simulations

Perform 1000 simulations, each with 40 samples of an exponential distribution.
The 40 samples will be used to calculate the arithmetic mean and variance and
then compared to the theoretical estimates.

To make the data reproducible, a seed will be set. Also, set the control 
parameters $\lambda = 0.2$ (the rate) and $n = 40$ (number of samples).

```{r exp-simulation, echo = TRUE}
# set seed for reproducability
set.seed(062000)

# set sampling values:
lambda <- 0.2             # rate parameter
n <- 40                   # number of samples (exponentials) in each simulation
numSimulations <- 1000    # number of simulations

# simulate the population
simMeans <- data.frame(expMean = sapply(1 : numSimulations, function(x) {mean(rexp(n, lambda))}))
```

## Sample Mean versus Theoretical Mean

According to the Central Limit Theorem, the distribution of the sample means
will be approximately normally distributed with a mean equal to the population
mean $\mu$ of the underlying distribution. Because the underlying distribution
in this simulation is exponential, the theoretical mean of the exponential
distribution will be compared to the corresponding sample mean of the
simulation. For an exponential distribution, the theoretical mean is equal to
$\frac{1}{\lambda}$. 

### Analysis

Calculate the sample mean and theoretical mean across all 1000 simulations of
40 samples from an exponential distribution where $\lambda = 0.2$. 

```{r sample-mean-vs-theoretical-mean, echo = TRUE}
# calculate sample mean and theoretical mean
sampleMean <- mean(simMeans$expMean)
theoMean <- 1/lambda
compMeans <- data.frame(sampleMean, theoMean)
names(compMeans) <- c("Sample Mean", "Theoretical Mean")
print(compMeans)
```

As part of the data analysis, also perform a one sample t-test to check the 95%
confidence interval for the sample mean.

```{r sample-mean-t-test, echo = TRUE}
t.test(simMeans$expMean, conf.level = 0.95)
```

### Plot Distribution

Display a histogram to show the averages of the 40 exponentials over 1000
simulations. Include the sample mean and theoretical mean for comparison.

```{r sample-mean-vs-theoretical-mean-plot, echo = TRUE}
# plot the distribution (sample mean versus theoretical mean)
expSimulationMeansChart <- ggplot(simMeans, aes(x = expMean, y = ..count..)) +
    geom_histogram(binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
    geom_vline(aes(xintercept = sampleMean, color = "sample"), size = 0.50) +
    geom_vline(aes(xintercept = theoMean, color = "theoretical"), size = 0.50) +
    xlab("Mean") +
    ylab("Frequency") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    scale_color_manual(name = "Means", values = c(sample = "blue", theoretical = "red")) +
    ggtitle("Distribution of Exponential Simulation Means")
print(expSimulationMeansChart)
```

### Findings

The sample mean came out to be `r sampleMean` while the theoretical mean is
`r theoMean`. As shown in the above chart, the mean of the sample means of
exponentials (blue vertical line) is very close to the theoretical mean of an
exponential distribution (red vertical line). We can also see that with a 95%
confidence interval, the sampled mean is between
`r t.test(simMeans$expMean, conf.level=0.95)$conf.int[1]` and
`r t.test(simMeans$expMean, conf.level=0.95)$conf.int[2]` which closely match.

## Sample Variance versus Theoretical Variance

In the same manner used to compare the
[Sample Mean and Theoretical Mean](#sample-mean-versus-theoretical-mean), the
Sample Variance will be compared to the Theoretical Variance.

### Analysis

The theoretical variance is $\frac{(\frac{1}{\lambda})^2}{n}$.

```{r sample-variance-vs-theoretical-variance, echo = TRUE}
# calculate sample variance and theoretical variance
sampleVariance <- var(simMeans$expMean)
theoVariance <- ((1/lambda)^2)/n
compVariance <- data.frame(sampleVariance, theoVariance)
names(compVariance) <- c("Sample Variance", "Theoretical Variance")
print(compVariance)
```

### Findings

The sample variance came out to be `r sampleVariance` which is very close
to the theoretical variance `r theoVariance`.

## Distribution

Determine whether the exponential distribution is approximately normally
distributed about the population mean. According to the Central Limit Theorem,
the means of the sample simulations should follow a normal distribution.

```{r distribution-of-exponential-means-plot, echo = TRUE}
# plot the distribution
expSimulationMeansChart <- ggplot(simMeans, aes(x = expMean)) +
    geom_histogram(aes(y = ..density..), binwidth = 0.15, color = "white", fill = rgb(0.2,0.7,0.1,0.4))  +
    geom_vline(aes(xintercept = sampleMean, color = "sample"), size = 0.50) +
    geom_vline(aes(xintercept = theoMean, color = "theoretical"), size = 0.50) +
    xlab("Mean") +
    ylab("Density") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    scale_color_manual(name = "Means", values = c(sample = "blue", theoretical = "red")) +
    stat_function(fun = dnorm, args = list(mean = sampleMean, sd = sqrt(sampleVariance)), color = "blue", size = 1.0) +
    stat_function(fun = dnorm, args = list(mean = theoMean, sd = sqrt(theoVariance)), color = "red", size = 1.0, linetype = "dashed") +
    ggtitle("Distribution of Exponential Simulation Means")
print(expSimulationMeansChart)
```

As shown in the above plot, the distribution of means of the sampled exponential
distribution appear to follow a normal distribution.

The density of the sampled data is shown by the light green bars. The dotted red
line represents a normal distribution which is very close to
the sample distribution colored in blue.
