---
title: "Statistical Inference Course Project (Part 2)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

## Overview

In this second part of a two-part project assignment, we are being asked to
investigate the effect of vitamin C on tooth growth in guinea pigs.

This analysis will compare the effects of different doses of vitamin C using
two delivery methods to study the tooth growth in guinea pigs. The sample
will consist of 60 guinea pigs. Each guinea pig will receive one of three dose
levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods,
orange juice or ascorbic acid.

The study will be performed using the ToothGrowth dataset which is included in
the R `datasets` package. The dataset is a data frame that contains 60
observations and 3 variables:

* len - A numeric vector indicating the measurement of tooth length after
        Vitamin C delivery
* supp - A factor vector describing the delivery method used. Either Orange
         Juice (OJ) or Ascorbic Acid (VC)
* dose - A numeric vector indicating the dosage level in milligrams
         (0.5, 1, or 2mg)

Further information on the ToothGrowth dataset can be found in the R
documentation using `?ToothGrowth`.

## Environment Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

Load packages used in this analysis.

```{r load-packages, echo = TRUE}
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}
if (!require(dplyr, warn.conflicts = FALSE)) {
    install.packages("dplyr", repos = "http://cran.us.r-project.org")
    library(dplyr, warn.conflicts = FALSE)
}
```

Load ToothGrowth dataset.

```{r load-data, echo = TRUE}
library(datasets)
data(ToothGrowth)
```

## Basic Data Summary

After loading the ToothGrowth dataset, provide a basic summary of the data.

```{r basic-data-summary, echo = TRUE}
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)

# unique observations grouped by delivery method and dosage level
ToothGrowth %>% select(supp, dose) %>% group_by(supp, dose) %>% summarize(count=n())

# summary of tooth length data grouped by delivery method and dosage level
by(data = ToothGrowth$len, INDICES = list(ToothGrowth$supp, ToothGrowth$dose), summary)
```

## Exploratory Data Analysis

Perform some basic exploratory data analyses of the data.

```{r exploratory-data-analysis, echo = TRUE}
```

## Confidence Intervals


## Hypothesis Test


## Conclusion


