---
title: "Statistical Inference Course Project (Part 2)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
  word_document:
    toc: yes
    df_print: paged
    keep_md: no
---

## Course Project

**Statistical Inference Course Project**

Peer-graded Assignment

* This course project is available on GitHub

    <a href="https://github.com/oraclejavanet/statistical-inference-course-project-1" target="_top">Statistical Inference Course Project</a>

## Synopsis

In this second part of a two-part project assignment, we are being asked to
investigate the effect of vitamin C on tooth growth in guinea pigs.

This analysis will compare the effects of different doses of vitamin C using
two delivery methods to study the tooth growth in guinea pigs. The sample
will consist of 60 guinea pigs. Each guinea pig will receive one of three dose
levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods,
orange juice or ascorbic acid supplements.

The study will be performed using the ToothGrowth dataset which is included in
the R `datasets` package. The dataset is a data frame that contains 60
observations and 3 variables:

* len - A numeric vector indicating the measurement of tooth length after
        vitamin C delivery
* supp - A factor vector describing the delivery method used: Orange Juice (OJ)
         or Ascorbic Acid (VC)
* dose - A numeric vector indicating the dosage level in milligrams per day
         (0.5, 1, or 2mg)

Further information on the ToothGrowth dataset can be found in the R
documentation using `?ToothGrowth`.

## Environment Setup

```{r setup, include = FALSE}
# set knitr options
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')

# clear global workspace (including hidden objects) for this session
rm(list = ls(all.names = TRUE))

# free up memory and display statistics on free memory
gc()

# disable scientific notation for numbers
options(scipen = 1)
```

Load packages used in this analysis.

```{r load-packages, echo = TRUE}
if (!require(ggplot2)) {
    install.packages("ggplot2")
    library(ggplot2)
}
```

Display session information.

```{r display-session-info, echo = TRUE}
sessionInfo()
```

## Basic Data Summary

After loading the ToothGrowth dataset, provide a basic summary of the data.

```{r basic-data-summary, echo = TRUE}
str(ToothGrowth)
summary(ToothGrowth)

# tabulate delivery method and dosage level values
table(ToothGrowth$supp, ToothGrowth$dose)

# summary of tooth length data grouped by delivery method and dosage level
by(data = ToothGrowth$len, INDICES = list(ToothGrowth$supp, ToothGrowth$dose), summary)
```

## Exploratory Data Analysis

Perform some basic exploratory data analyses of the data. The analyses will
explore the following relations:

1. Tooth Length (len) as a function of Delivery Method (supp)
1. Tooth Length (len) as a function of Dosage Level (dose)
1. Tooth Length (len) as a function of Delivery Method (supp) and Dosage Level (dose)

### Tooth Length to Delivery Method

```{r tooth-length-relative-to-delivery-method, echo = TRUE}
tg <- ToothGrowth
levels(tg$supp) <- c("Orange Juice", "Ascorbic Acid")
gLenSupp <- ggplot(data = tg, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Delivery Method")
print(gLenSupp)
```

**Observation**

The above chart shows that using orange juice as the delivery method,
independent of dosage level, had a more favorable effect on tooth growth than
ascorbic acid.

### Tooth Length to Dosage Level

```{r tooth-length-relative-to-dosage-level, echo = TRUE}
gLenDose <- ggplot(data = ToothGrowth, aes(x = factor(dose), y = len)) + 
    geom_boxplot(aes(fill = factor(dose))) +
    xlab("Dosage Level (mg/day)") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="dose")) +
    theme(plot.title = element_text(size = 14, hjust = 0.5)) +
    ggtitle("Tooth Length as a Function of Dosage Level")
print(gLenDose)
```

**Observation**

The above chart shows a positive relationship that higher dosage levels of
vitamin C, independent of delivery method, had a more favorable effect on tooth
growth than lower dosages of vitamin C. The 2 mg/day dosage level had the
best effect on tooth growth followed by 1 mg/day and then 0.5 mg/day.

### Tooth Length to Delivery Method and Dosage Level

```{r tooth-length-relative-to-delivery-method-and-dosage-level, echo = TRUE}
tg <- ToothGrowth
levels(tg$supp) <- c("Orange Juice", "Ascorbic Acid")
gLenSuppDose <- ggplot(data = tg, aes(x = supp, y = len)) + 
    geom_boxplot(aes(fill = supp)) +
    facet_wrap(~ dose) +
    xlab("Delivery Method") +
    ylab("Tooth Length") +
    guides(fill=guide_legend(title="supp")) +
    theme(plot.title = element_text(size = 14, hjust = 0.5, vjust = 0.5),
          axis.text.x = element_text(angle = 45,
                                     hjust = 0.5,
                                     vjust = 0.5,
                                     margin = margin(b = 10))) +
    ggtitle("Tooth Length as a Function of\nDelivery Method and Dosage Level")
print(gLenSuppDose)
```

**Observation**

Looking at the above chart which shows tooth growth as a function of delivery
method and dosage level, it appears that orange juice is more effective than
ascorbic acid as the delivery method when the dosage level is 0.5 to 1 mg/day.
The higher dosage level of 2 mg/day is more effective than the lower
dosages; however, both delivery methods are equally as effective.

## Inferential Statistics

Exploratory data analysis is helpful to study and visually interrogate data;
however, the use of inferential statistics allows us to use techniques such as
the confidence interval and/or hypothesis testing to more precisely draw
conclusions about a population from which representative samples were taken.

In this section, hypothesis testing will be employed to study the impact of
delivery method (orange juice and ascorbic acid supplements) and dosage level
on tooth growth. A p-value less than or greater than a significance level of 5%
will be used as the threshold to reject or accept the null hypothesis.

### Hypothesis 1

Hypothesis testing will be conducted to study the impact of delivery method on
tooth growth, independent of dosage level. A t-test will be performed on the
null hypothesis that the two delivery methods have no effect on
tooth growth.

```{r inferential-statistics-test-1, echo = TRUE}
t1 <- t.test(len ~ supp, data = ToothGrowth, conf.level = 0.95)
paste0("p-value = ", round(t1$p.value, 4))
paste0("confidence interval = (", round(t1$conf.int[1], 4) , ", ", round(t1$conf.int[2], 4), ")")
```

The observed p-value `r round(t1$p.value, 4)` is greater than 0.05 and the
95% confidence interval contains zero. This indicates weak evidence against the
null hypothesis so we fail to reject the null hypothesis.

### Hypothesis 2

Hypothesis testing will be conducted to study the impact of delivery method on
tooth growth for a single dosage level. A t-test will be performed on the null
hypothesis that the two delivery methods at a dosage level of
0.5 mg/day have no effect on tooth growth.

```{r inferential-statistics-test-2, echo = TRUE}
t2 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5), conf.level = 0.95)
paste0("p-value = ", round(t2$p.value, 4))
paste0("confidence interval = (", round(t2$conf.int[1], 4) , ", ", round(t2$conf.int[2], 4), ")")
```

The observed p-value `r round(t2$p.value, 4)` is less than 0.05 and the
95% confidence interval does not contain zero. This indicates strong evidence
against the null hypothesis so the null hypothesis can be rejected.

### Hypothesis 3

Hypothesis testing will be conducted to study the impact of delivery method on
tooth growth for a single dosage level. A t-test will be performed on the null
hypothesis that the two delivery methods at a dosage level of 1 mg/day have no
effect on tooth growth.

```{r inferential-statistics-test-3, echo = TRUE}
t3 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 1), conf.level = 0.95)
paste0("p-value = ", round(t3$p.value, 4))
paste0("confidence interval = (", round(t3$conf.int[1], 4) , ", ", round(t3$conf.int[2], 4), ")")
```

The observed p-value `r round(t3$p.value, 4)` is less than 0.05 and the
95% confidence interval does not contain zero. This indicates strong evidence
against the null hypothesis so the null hypothesis can be rejected.

### Hypothesis 4

Hypothesis testing will be conducted to study the impact of delivery method on
tooth growth for a single dosage level. A t-test will be performed on the null
hypothesis that the two delivery methods at a dosage level of 2 mg/day have no
effect on tooth growth.

```{r inferential-statistics-test-4, echo = TRUE}
t4 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 2), conf.level = 0.95)
paste0("p-value = ", round(t4$p.value, 4))
paste0("confidence interval = (", round(t4$conf.int[1], 4) , ", ", round(t4$conf.int[2], 4), ")")
```

The observed p-value `r round(t4$p.value, 4)` is greater than 0.05 and the
95% confidence interval contains zero. This indicates weak evidence against the
null hypothesis so we fail to reject the null hypothesis.

## Conclusion

This analysis employed the use of statistical inference to compare the effects
of different dosage levels of vitamin C using two delivery methods to study the
tooth growth in guinea pigs. 

### Assumptions

* Tooth growth follows a normal distribution.

* The two variables, dosage levels and delivery method, are independent and
  identically distributed.

* No other confounding factors were included that would effect tooth growth.

* The population was comprised of similar guinea pigs.

* A p-value less than or greater than a significance level of 5% will be used as
  the threshold to reject or accept the null hypothesis.

### Conclusions

Based on exploratory data analysis and confirmed by hypothesis tests and
associated confidence intervals, we can safely infer that an increase in
dosage levels of vitamin C increases tooth growth. However, when studying tooth
growth as a function of both dosage level and delivery method, it is
inconclusive whether or not the delivery method had any effect on tooth growth.
While it started to appear that orange juice was more effective than ascorbic
acid, this only occurred at dosage levels of 0.5 to 1 mg/day. At 2 mg/day, both
delivery methods were equally as effective.
