---
title: "Statistical Inference Course Project (Part 2)"
author: "Jeffrey M. Hunter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: cosmo
    keep_md: no
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='figures/')
```

## Overview

In this second part of a two-part project assignment, we are being asked to
investigate the effect of vitamin C on tooth growth in guinea pigs.

This analysis will compare the effects of different doses of vitamin C using
two delivery methods to study the tooth growth in guinea pigs. The sample
will consist of 60 guinea pigs. Each guinea pig will receive one of three dose
levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods,
orange juice or ascorbic acid.

The study will be performed using the ToothGrowth dataset which is included in
the R `datasets` package. The dataset is a data frame that contains 60
observations and 3 variables:

* len - A numeric vector indicating the measurement of tooth length after
        Vitamin C delivery
* supp - A factor vector describing the delivery method used. Either Orange
         Juice (OJ) or Ascorbic Acid (VC)
* dose - A numeric vector indicating the dosage level in milligrams
         (0.5, 1, or 2mg)

Further information on the ToothGrowth dataset can be found in the R
documentation (?ToothGrowth).

## Load Packages

Start by loading packages used in this analysis.

```{r load-packages, echo = TRUE}
if (!require(ggplot2)) {
    install.packages("ggplot2", repos = "http://cran.us.r-project.org")
    library(ggplot2)
}
```



## Exploratory Data Analysis






```{r exploratory-data-analysis, echo = TRUE}
# load ToothGrowth dataset
library(datasets)
data(ToothGrowth)
```

## Basic Summary

Load the ToothGrowth data and perform some basic exploratory data analyses Provide a basic summary of the data.

## Confidence Intervals


## Hypothesis Test


## Conclusion

Following results can be concluded from the above hypothesis testing

* Supplement type has no effect on tooth growth.

* Increasing the dose level leads to increased tooth growth.

===============

Both the confidence intervals and hypothesis tests indicate that at lower dosages, specifically 0.5 mg and 1 mg, the Orange Juice delivery method results in longer teeth than the Ascorbic Acid. However, at higher dosages, specifically 2 mg, there appears to be little difference and certainly not enough to reject the null hypothesis that they have the same effect.

However, the plot of means and confidence intervals does hint at a flattening out of effectiveness when using the Orange Juice method at a faster rate than when using the Ascorbic Acid method. This might warrant further exploration to see if Ascorbic Acid results in greater tooth length than Orange Juice and higher dosages.
